Classical Simulation of Quantum Error Correction in a Fibonacci Anyon Code

### 2015-06-03 fib2.tex
deleted leftover working from diagram making
revised fusion diagram thing, made it into a figure completed discussion of simulation added new jw00t ref deleted comments in tex file other minor edits  - its now a couple paragraphs too long, but that should be fairly easy to fix.

\begin{abstract}
Classically simulating the dynamics of excitations in two-dimensional quantum systems that 
support anyons is likely intractable in general because such dynamics are sufficient to implement 
universal quantum computation. However, processes of interest for the study of quantum 
error correction in anyon systems are typically drawn from a restricted class that displays 
significant structure over a wide range of system parameters.
We exploit this structure to classically simulate, and thereby demonstrate the success of, an 
error-correction protocol for a quantum memory based on the universal Fibonacci anyon 
model.  We numerically simulate a phenomenological model of the system and noise 
processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in the regimes where we expect successful error correction to be possible. This insight allows us to simulate the noise and recovery processes for a quantum code based on a universal anyon model.

@@

The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
	\begin{align}
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;R^{ab}_c\cdot\;
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}
where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, denoted as
	\begin{align}
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(-1,2);
			\draw(2,3)--(0,1);
			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(1,2);
			\draw(2,3)--(0,1);
			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Our simulation is in effect that of a $(2+1)$-dimensional topological quantum field theory, which allows for topological processes such as pair-creation, braiding, and fusion of anyonic charges. This is modelled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside.  For more details of an analogous model, see~\cite{Brell2013}.

As was noted in~\cite{Pfeifer2014}, when dealing with two-dimensional 
anyon dynamics, it is not sufficient to consider only
the algebraic (fusion/braiding) properties of the anyons,
one must also account for the algebra of observables.

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\mathbb{I}}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathbb{I}$ at site $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

In order to perform a logical error on our code, a noise process must have support on a 
ribbon running around a homologically nontrivial loop of the lattice. These correspond to 
processes in which anyonic charge is transported around the non-trivial loop before 
annihilating to vacuum.
% cggb{we want to define logical operators as those that map the codespace to itself, so for it to be a logical operator, they must annihilate to vacuum. they need not be unitary}.

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
%or the clustering length scale exceeds the size of the lattice.
%\cggb{our simulation may terminate when the decoder performs a nontrivial operation, but here we're just describing the decoder, which can't know that.}
%\dude{in our sim the decoder stops when there is one cluster. Ie., it never gives up.}
%\cggb{A few more details? What is our length schedule?} 
%\cggb{How does our decoder relate to other decoding schemes used in~\cite{Wootton2013,Brell2013,Hutter2014}?} 
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

@@

Of course, if we were to consider noise so strong that 
the components percolate over the entire lattice, we would no 
longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Below the point at which the decoder begins to create such nontrivial loops, our simulation remains efficient,
%\dude{efficiency has nothing to do with it}
while it begins to declare failure whenever non-trivial logical 
operations occur. This allows us to simulate large lattice sizes 
in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

Our concrete simulation proceeds by associating a tile with each site of the 
$L\times L$ periodic square lattice.
In order to keep track of the observables of the system 
we maintain a set of disjoint curves, representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. 
In general, the curves wind around the two 
dimensional tiles in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding 
to disjoint components.
For comparison, previous work 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}
\caption{Four equivalent representations of a state,
along with two syndrome observables marked with solid and dashed lines:
(a) two pairs are joined in (b).
Deform the curve straight to get (c),
then effect some
braid operations on the state so that (d) the two
observables are now represented on the curve. 
}
\label{f:syndrome}
\end{center}
\end{figure}

In \Fref{f:syndrome} we show two adjacent sites
involved in a syndrome measurement.
%This will be the total charge on each site, ie. either $\tau$ or $\mathbb{I}.$
The error process has resulted in two pair
creation events along the edge joining these sites.
\dude{i'd like to write out the equations for each of
these 4 steps, but we'd need to keep it tight somehow.}

\newcommand{\verteq}{\rotatebox{90}{$\,=$}}
\begin{align*}
		&\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(1,0)--(0,1)[dashed];
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
			\node at (1,3) [anchor=west] {{\scriptsize $2$}};
			%			
			\begin{scope}[xshift=4cm]
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $3$}};
				\node at (1,3) [anchor=west] {{\scriptsize $4$}};
			\end{scope}
		\end{tikzpicture}\\
		&\qquad\;\downarrow\\
		&\begin{tikzpicture}[baseline=0.5cm,scale=0.3]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
			\end{scope}
			%			
			\begin{scope}[xshift=4cm]
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(1,4)--(1,5);
				\draw(0,1)--(3,4)--(3,5);
				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
			\end{scope}
		\end{tikzpicture}\\
		&\qquad\downarrow\\
		&\begin{tikzpicture}[baseline=1.5cm,scale=0.3]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,7);
				\draw(0,1)--(1,2)--(1,5);
				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
				\node at (-1,7) [anchor=west] {{\scriptsize $1$}};
				\node at (1,7) [anchor=west] {{\scriptsize $4$}};
			\end{scope}
			%			
			\begin{scope}[xshift=4cm]
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(1,4)--(1,5);
				\draw(0,1)--(3,4)--(3,5);
				\draw(3,7)--(3,9);
				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
				\node at (1,9) [anchor=west] {{\scriptsize $2$}};
				\node at (3,9) [anchor=west] {{\scriptsize $3$}};
			\end{scope}
			%
			\begin{scope}[xshift=6cm,yshift=2cm]
				\draw plot [smooth, tension=1] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
				\draw plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
			\end{scope}
			\begin{scope}[xshift=4cm,yshift=4cm]
				\draw plot [smooth, tension=1] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
				\draw plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
			\end{scope}
		\end{tikzpicture}\\
		&\qquad\verteq\\
		\sum_{j,j'\in\{\mathbb{I},\tau\}}R^{\tau\tau}_{\mathbb{I}}[F^{\tau\tau\tau}_{\tau}]_{\mathbb{I}j}R^{\tau\tau}_{j}[F^{\tau\tau\tau}_{\tau}]^{-1}_{jj'}\;\;
		&\begin{tikzpicture}[baseline=0.8cm,scale=0.3]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dotted];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
				\node at (1,3) [anchor=west] {{\scriptsize $4$}};
				\node at (0.7,0.2) [anchor=east] {{\scriptsize $j'$}};
			\end{scope}
			%			
			\begin{scope}[xshift=4cm]
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(1,4)--(1,5);
				\draw(0,1)--(3,4)--(3,5);
				\node at (1,5) [anchor=west] {{\scriptsize $2$}};
				\node at (3,5) [anchor=west] {{\scriptsize $3$}};
			\end{scope}
		\end{tikzpicture}\\
		&\qquad\verteq\\
		\sum_{j,j'\in\{\mathbb{I},\tau\}}R^{\tau\tau}_{\mathbb{I}}[F^{\tau\tau\tau}_{\tau}]_{\mathbb{I}j}R^{\tau\tau}_{j}[F^{\tau\tau\tau}_{\tau}]^{-1}_{jj'}\;\;
		&\begin{tikzpicture}[baseline=0.2cm,scale=0.3]		
			\draw(2,-1)--(0,1)[dotted];
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
			\node at (1,3) [anchor=west] {{\scriptsize $4$}};
			\node at (1,0) [anchor=east] {{\scriptsize $j'$}};
			\draw(2,-1)--(2,-2)[dashed];
			%			
			\begin{scope}[xshift=4cm]
				\draw(-2,-1)--(0,1)[dotted];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $2$}};
				\node at (1,3) [anchor=west] {{\scriptsize $3$}};
				\node at (-1,0) [anchor=west] {{\scriptsize $\bar{j'}$}};
			\end{scope}
		\end{tikzpicture}
\end{align*}
\cggb{Here 1,2,3,4 are meant to refer to the 4 different leaves in \Fref{f:syndrome} a) taken clockwise from the top left (equivalently those in c) taken down the page). The first diagram represents a), the second b) and c) together, and the last three d). Is this correct? Of course I would clean and explain these properly if this is the kind of thing we want. Also I haven't been very careful with the calculation. I would need to check if there is a Frobenius-Schur indicator missing on the last line. We could also do something like make circles above 1,4 and 2,3 at the end to show the measurements being made in the figure or something.}

>>>>

\begin{abstract}
Classically simulating the dynamics of anyonic excitations in two-dimensional quantum systems is likely intractable in general because such dynamics are sufficient to implement 
universal quantum computation. However, processes of interest for the study of quantum 
error correction in anyon systems are typically drawn from a restricted class that displays 
significant structure over a wide range of system parameters.
We exploit this structure to classically simulate, and thereby demonstrate the success of, an 
error-correction protocol for a quantum memory based on the universal Fibonacci anyon 
model.  We numerically simulate a phenomenological model of the system and noise 
processes on lattice sizes of up to 
$128\times128$ sites, and find a lower bound on the error-correction threshold of 
approximately $12.5\%$, which is comparable to those previously known for abelian and 
(non-universal) nonabelian anyon models.
\end{abstract}

@@

Recent investigations have begun to explore quantum error correction for nonabelian anyon 
models~\cite{Brell2013, Wootton2013, Hutter2014, Wootton2015}. Nonabelian anyon models are especially interesting 
because braiding and fusion of these anyons in general allows for the implementation of universal quantum 
computation. However, the initial studies of error-correction in nonabelian anyon systems have focused on specific models, such as the Ising 
anyons~\cite{Brell2013} and the so-called $\Phi$-$\Lambda$ 
model~\cite{Wootton2013, Hutter2014} that, while nonabelian, are not universal. The general dynamics of these particular anyon models is known to be efficiently classically simulable, a fact
that was exploited to enable efficient simulation of error correction in these systems. When considering more general anyon models, their ability to perform universal quantum computation would seem a significant barrier to their simulation on a classical computer. While simulation of general dynamics does indeed seem intractable, we argue that the kinds of processes that are typical of thermal noise are sufficiently structured to allow for their classical simulation in the regimes where we expect successful error correction to be possible. This insight allows us to simulate the noise and recovery processes for a quantum code based on a universal anyon model.

@@

The dynamics and history of a system of anyons can be conveniently represented by a graphical calculus, which can be regarded as a representation of the world-lines of the particles. These diagrams can be manipulated by $R$-moves, corresponding to braiding processes:
	\begin{align}
	\label{e:rmove}
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw plot [smooth, tension=1] coordinates { (0,1) (0.75,1.8) (-0.6,2.45) (-1,3)};
			\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\draw plot [smooth, tension=1] coordinates { (0,1) (-0.75,1.8) (0.6,2.45) (1,3)};
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;R^{ab}_c\cdot\;
		\begin{tikzpicture}[baseline=0.5cm,scale=0.3]		
			\draw(0,0)--(0,1);
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (0,0) [anchor=west] {{\scriptsize $c$}};
			\node at (-1,3) [anchor=east] {{\scriptsize $b$}};
			\node at (1,3) [anchor=west] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}
where $a$, $b$, and $c$ represent particle types, or $F$-moves, representing the associativity rule for fusion processes, denoted as
	\begin{align}
		\label{e:fmove}
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(-1,2);
			\draw(2,3)--(0,1);
			\node at (-0.45,1.6) [anchor=north east] {{\scriptsize $i$}};
			\node at (0,0) [anchor=west] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
		&\;\;=\;\;\sum_j[F^{abc}_d]_{ij}\cdot\;
		\begin{tikzpicture}[baseline=0.6cm,scale=0.3]		
			\draw(0,0)--(0,1)--(-2,3);
			\draw(0,3)--(1,2);
			\draw(2,3)--(0,1);
			\node at (0.42,1.63) [anchor=north west] {{\scriptsize $j$}};
			\node at (0,0) [anchor=east] {{\scriptsize $d$}};
			\node at (2,3) [anchor=east] {{\scriptsize $c$}};
			\node at (0,3) [anchor=east] {{\scriptsize $b$}};
			\node at (-2,3) [anchor=east] {{\scriptsize $a$}};
		\end{tikzpicture}
	\end{align}

@@

We use a phenomenological model of Fibonacci anyon dynamics which neglects any 
microscopic details of the system. 
This is consistent with the principles of topologically 
ordered systems and anyonic physics, where the key universal features describing the 
anyon model correspond to large length-scale physics, while the microscopic physics plays 
a less important (and non-universal) role.
Our simulation is in effect that of a $(2+1)$-dimensional topological quantum field theory, which allows for topological processes such as pair-creation, braiding, and fusion of anyonic charges. This is modelled on a periodic $L\times L$ square lattice $\Lambda$ of \emph{tiles} on which anyons can reside.  This treatment applies equally well to lattice models, where anyons can only be located at a discrete set of points, or continuum models, where a discretization of the manifold must be chosen in order to perform charge measurements. In the former case, each tile of $\Lambda$ represents a lattice site, while in the latter case each tile represents a measurement region. For more details of an analogous model, see~\cite{Brell2013}.

As was noted in~\cite{Pfeifer2014}, when dealing with two-dimensional 
anyon dynamics, it is not sufficient to consider only
the algebraic (fusion/braiding) properties of the anyons,
one must also account for the algebra of observables. \cggb{This sentence seems odd here. Consider moving?}

@@

We consider encoding a qubit of quantum information in the global degeneracy associated 
with the topology of the manifold of our system. We imagine an idealized Hamiltonian for this 
system of the form
\begin{align}
	H=-\sum_{i\in \Lambda}\proj{\mathbb{I}}_i\;,\label{e:hamiltonian}
\end{align}
with $\proj{\mathbb{I}}_i$ the projector to charge $\mathbb{I}$ at site (or tile) $i$, i.e.~the ground 
space of the model has vacuum (or no anyon) at each site. Given the toroidal boundary 
conditions, this space is two-fold degenerate as noted above. We identify this ground space 
as the codespace of our model. Typical realistic noise processes in this kind of system are 
pair-creation, hopping, exchange etc.~of anyons on neighboring sites of $\Lambda$. It was 
seen in~\cite{Brell2013} that a simplified noise model consisting of pair-creation events only 
is sufficient to capture the qualitative features of an error-correction simulation, and so for convenience we 
will restrict to pair-creation noise processes in the numerical results presented in this study. 
Such pair-creation events act on neighboring pairs of sites,
and so may be associated with an edge of the lattice, chosen 
uniformly at random at each time step of the simulation. The simulation time, and thus the 
error-correction threshold, will be measured in terms of average number of noise processes 
per edge, as opposed to an iid noise probability per edge. The latter measure is not 
appropriate for error processes in nonabelian anyon models, but the two coincide in the limit 
of low error rates in those cases where they are comparable (see Ref.~\cite{Brell2013}).

In order to perform a logical error on our code, a noise process must have support on a 
ribbon running around a homologically nontrivial loop of the lattice. These correspond to 
processes in which anyonic charge is transported around the non-trivial loop before 
annihilating to vacuum.

Our error-correction algorithm is
based on a hierarchical clustering algorithm~\cite{Hastie2009, Wootton2015},
and follows a similar strategy to the hard-decision renormalization group decoder~\cite{Bravyi2011}. 
The decoder proceeds by measuring the occupations of 
anyons at each site of the lattice, producing a \emph{syndrome} of occupied sites. Following this, it forms clusters of nearby anyons and 
measures the total charge within each cluster.
Clusters with trivial charge are discarded, and
the rest are joined (agglomeratively~\cite{Hastie2009})
at linearly increasing length scales. This iterative process concludes when there is at most a single cluster remaining (see \Fref{f:decode}).
Note that unlike the case of abelian anyons, the decoder
cannot determine the charge of each cluster given only the syndrome information (as in \cite{Bravyi2011}),
and so will query (measure) the system to find this.
Our simulation proceeds in this way as a dialogue between decoder
and system, terminating when either the decoder itself terminates or when it performs a homologically nontrivial operation, thereby altering the codespace (and failing to perform successful information recovery).
Note that, as was found in Ref.~\cite{Brell2013}, we expect that our qualitative results may be reproduced by most alternative families of decoders, although it may not be clear in all cases how to extend these decoders to the non-abelian setting.
The advantage of using a clustering decoder is 
its simplicity and flexibility, and the fact that its clustering scheme is compatible with the structure in the noise processes that allows us to classically simulate them.

@@

Of course, if we were to consider noise so strong that 
the components percolate over the entire lattice, we would no 
longer expect to be able to efficiently simulate the noise and recovery processes in the system. 
However, recall that logical errors in our system correspond to processes that act over a non-trivial loop of the torus. 
Below the point at which the decoder begins to create such nontrivial loops, our simulation remains efficient,
while it begins to declare failure whenever non-trivial logical 
operations occur. This allows us to simulate large lattice sizes 
in the successful error-correction (sub-percolated) regime, including up to $L=128$.

@@

%Our concrete simulation proceeds by associating a tile with each site of the 
%$L\times L$ periodic square lattice $\Lambda$.

In order to keep track of the state of the system 
we maintain a set of disjoint curves, \cggb{do they need to be directed, or is that just for the figure?} representing the history of anyons on the lattice.
These curves can also be thought of as a top-down view of the fusion tree:
the anyons comprising the fusion tree are here displayed as points along each curve. 
In general, the curves wind around our two 
dimensional manifold in a haphazard way determined by the progress of the simulation.
This allows a dynamically generated basis for the fusion space that easily allows for tensor 
decomposition into factors corresponding 
to disjoint components. 
used a single fixed basis for the fusion space (determined by row-major ordering of 
sites)~\cite{Brell2013}.

The application of a pair-creation noise process corresponds to the addition of an extra piece of curve to the lattice, along with two preferred points on it denoting the anyon locations. Following the noise processes, our simulation must proceed to measure the total charge within each tile (the results of these measurements will form the error syndrome). This requires joining any curves within that tile, and performing braid moves along the curves until all anyons within the tile have been localized within a contiguous region of the curve. Following this, the total charge can (probabilistically) simply be read off the relevant section of the fusion tree.
An example of this procedure for a simple noise process is shown in \Fref{f:syndrome}, and the corresponding braid moves are shown in \Fref{f:syndrome2}. The technical details of the algorithm and data structures used
for this procedure will be discussed elsewhere.
Braiding processes that encircle a non-trivial loop of our (toric) manifold can also be treated in an analogous way, following e.g.~\cite{Pfeifer2012}.
\cggb{This paragraph and the next still needs another look.}

\begin{figure}[th!]
\begin{center}
	\includegraphics[width=1.0\columnwidth]{pic-syndrome.pdf}\\
\caption{(a) Noise processes initially form isolated sets of pair-created anyons, each crossing the boundary of a tile. These events are each represented by curves, where anyons are denoted by specific points on these curves. (b) In order to measure the total charge contained within each tile, we must first join the fusion trees associated with all such pairs. This corresponds to joining the edges associated with each pair into a continuous path, representing a basis for the fusion space chosen arbitrarily. (c) The total charge of each tile can then be found by braiding anyons around each other until all charges within a tile are neighbours on the path, as in (d). \cggb{We should think about the best way to represent the transition from (c) to (d). I think the current red lines might be confusing, but I don't have any great ideas.}}
\label{f:syndrome}
\end{center}
\end{figure}

\begin{figure}[th!]
\begin{center}
			\begin{tikzpicture}[baseline=0.7cm,scale=0.3, xscale=0.9]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
			\end{scope}
			%			
			\begin{scope}[xshift=4cm]
				\draw(0,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(1,4)--(1,5);
				\draw(0,1)--(3,4)--(3,5);
				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
			\end{scope}
		\end{tikzpicture}
		\hspace{-3mm}$\longrightarrow$\hspace{1mm}
		\begin{tikzpicture}[baseline=1.5cm,scale=0.3, xscale=1]
			\begin{scope}[xshift=2cm,yshift=2cm]		
				\draw(1,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(-1,7);
				\draw(0,1)--(1,2)--(1,5);
				\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
				\node at (1,3) [anchor=west] {{\scriptsize $2$}};
				\node at (-1,7) [anchor=west] {{\scriptsize $1$}};
				\node at (1,7) [anchor=west] {{\scriptsize $4$}};
			\end{scope}
			%			
			\begin{scope}[xshift=4cm]
				\draw(0,0)--(0,1)[dashed];
				\draw(0,1)--(-1,2)--(1,4)--(1,5);
				\draw(0,1)--(3,4)--(3,5);
				\draw(3,7)--(3,9);
				\node at (1,5) [anchor=west] {{\scriptsize $3$}};
				\node at (3,5) [anchor=west] {{\scriptsize $4$}};
				\node at (1,9) [anchor=west] {{\scriptsize $2$}};
				\node at (3,9) [anchor=west] {{\scriptsize $3$}};
			\end{scope}
			%
			\begin{scope}[xshift=6cm,yshift=2cm]
				\draw plot [smooth, tension=1] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
				\draw plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
			\end{scope}
			\begin{scope}[xshift=4cm,yshift=4cm]
				\draw plot [smooth, tension=1] coordinates { (1,3) (0.75,3.8) (-0.6,4.45) (-1,5)};
				\draw[color=white, line width=7] plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
				\draw plot [smooth, tension=1] coordinates { (-1,3) (-0.75,3.8) (0.6,4.45) (1,5)};
			\end{scope}
		\end{tikzpicture}
		$\xrightarrow{R, F}$
		\begin{tikzpicture}[baseline=0.2cm,scale=0.3,xscale=0.9]		
			\draw(2,-1)--(0,1)[];
			\draw(0,1)--(-1,2)--(-1,3);
			\draw(0,1)--(1,2)--(1,3);
			\node at (-1,3) [anchor=west] {{\scriptsize $1$}};
			\node at (1,3) [anchor=west] {{\scriptsize $4$}};
			\node at (1,-0.05) [anchor=east] {{\scriptsize $q$}};
			\draw(2,-1)--(2,-2)[dashed];
			%			
			\begin{scope}[xshift=4cm]
				\draw(-2,-1)--(0,1)[];
				\draw(0,1)--(-1,2)--(-1,3);
				\draw(0,1)--(1,2)--(1,3);
				\node at (-1,3) [anchor=west] {{\scriptsize $2$}};
				\node at (1,3) [anchor=west] {{\scriptsize $3$}};
				\node at (-1,0) [anchor=west] {{\scriptsize $q'$}};
			\end{scope}
		\end{tikzpicture}
\caption{In order to transform from the configuration shown in \Fref{f:syndrome}(c) to \ref{f:syndrome}(d), several operations must be performed. The labels of the four leaves of these fusion trees correspond to the four anyons taken from the top of the path in \Fref{f:syndrome}(c) to the bottom. Vacuum charges are denoted by dashed lines. In order to rearrange the anyons so that particles 1 and 4 are neighbors, several braid moves must be performed as shown. The effect of these on the fusion tree can then be evaluated by the application of several $R$ and $F$ moves as defined in equations (\ref{e:rmove}) and (\ref{e:fmove}). After the manipulations shown, the total charges within each tile can be read off the diagram as $q$ and $q'$.}
\label{f:syndrome2}
\end{center}
\end{figure}

Following the result of these charge measurements, the decoder determines a recovery operation that involves fusion of subsets of anyons. These fusion results can be calculated in an analogous way, and the output charge placed at an appropriate point in the lattice. This procedure is iterated until either the decoder terminates successfully, or the simulation itself declares failure.
